Improving Usability of Remote Access to Sensitive Data

### 2014-07-23 ccsw14/paper.tex
finished design section

The virtual machine infrastructure that underlies the Data Capsules system is
trusted and assumed to be well implemented.  While such an assumption may be
somewhat controversial, we note its use in prior work~\cite{capsules}, and also
note the existence of microkernels such as seL4 and OKL4~\cite{sel4, OKL4},
which serve as a proof of concept that it is possible to create a lightweight
hypervisor with a reasonable degree of trustworthiness.  However, the
development of such a hypervisor lies outside the scope of this paper.

@@

The fundamental element that provides the guarantees

@@

>>>>

The virtual machine infrastructure that underlies the Data Capsules system is
trusted and assumed to be well implemented.  While this assumption may be
somewhat controversial, we note its use in prior work~\cite{capsules}, and also
note the existence of microkernels such as seL4 and OKL4~\cite{sel4, OKL4},
which serve as a proof of concept that it is possible to create a lightweight
hypervisor with a reasonable degree of trustworthiness based in formal proofs
of correctness.  However, the development of such a hypervisor lies outside the
scope of this paper.

@@

The major element that is used to isolate VMs that are in the middle of
accessing sensitive data is live migration \cite{migration-clark}.  When a VM
switches from maintenance mode to secure mode, the underlying virtual machine
manager migrates that VM to a separate physical machine that has access to the
sensitive data, but which is configured to lack network access for any purpose
other than to receive other incoming VM migrations.  In other words, VMs in
maintenance mode are strictly segregated within separate physical machines from
VMs running in secure mode.  We can therefore refer to a physical machine
running only maintenance VMs as a \emph{maintenance machine}, and likewise a
\emph{secure machine} runs only secure mode instantiations of VMs.

By adopting this segregated design, the threat of data leak via covert channels
within a single physical machine is eliminated, so long as the trusted computing
base remains uncompromised.  The remaining risk is that of one VM in secure mode
leaking data to another VM that is also in secure mode, but there is no reason
to want to do so, as both VMs would already have access to the entire secure
repository.  On a machine running VMs in maintenance mode, there is no way to
access the sensitive data, and hence no data to be leaked.  We stress that this
design also requires at least two physical machines in order to provide these
security guarantees.

@@

As the name suggests, maintenance mode is intended to allow a user to maintain
their VM's software environment, performing actions such as updating software
and uploading any files needed to conduct their research when in secure mode.
In accordance with these needs, maintenance mode allows the user full Internet
access and persistent storage (i.e.  changes to the virtual hard disk will
persist on reboot) in exchange for a lack of access to the sensitive data and
the user's persistent secure storage volume.

When a user initiates the switch to secure mode, their system state is migrated
to a secure machine.  Once this migration has completed, the state of the
maintenance VM is frozen, and remains in this frozen state until the user exits
secure mode.  Upon exit from secure mode, the maintenance VM is simply unfrozen,
and continues processing from the point at which it left off at the time of the
original switch to secure mode.

\subsection{Secure Mode}

When the user switches to secure mode, the state of their VM, previously in
maintenance mode, is migrated to a secure machine where it continues execution.
By switching to secure mode, the user sacrifices network access and virtual hard
disk storage persistence in exchange for access to the sensitive data and their
secure storage volume, which is mounted upon entering secure mode.  Any changes
to the normal storage are stored only in the machine's virtual memory, and
discarded when the user exits secure mode.  Only changes to the user's secure
storage volume will persist to the next time the user enters secure mode.
Indeed, when the user switches back to maintenance mode from secure mode, all
accumulated state change after the original switch into secure mode is discarded
from memory, and the maintenance VM simply resumes execution from where it left
off prior to the switch.

Conceptually, we can also think of this functionality in this way: by using the
same single virtual machine and transporting it, via migration, to a different
environment where it is subject to a different set of permissions and
privileges, we can allow the user to transport a custom software environment
created in an insecure environment over to a secure environment where it can
then be used to conduct research, and subsequently be discarded.  By giving them
a secure storage volume in secure mode, the user can save their intermediate
results before further iterating their software environment, preserving
flexibility in the case that they require further enhancements to their
software environment.

The use of two separate modes for a single virtual machine, as opposed to, say,
using two separate machines or two separate VMs, allows the user to customize
the software environment they use when the secure data is accessed.  This would
not be possible without opening high-risk channels or a burdensome and
inconvenient auditing process under a two (virtual) machine model.
